---
title: "Assignment 9"
output: pdf_document
author: "Natalie Holsclaw"
date: "2025-03-23"
editor_options: 
  chunk_output_type: console
---

### (1) Logistic Regression

```{r}
# Load necessary packages
library(here)
library(tidyverse)
library(naniar)
library(DHARMa)
library(gtsummary)
library(broom)

# Load in data
lizard <- read_csv("jrn_lizard.csv")

# filter data
sb_lizard <- lizard %>% 
  filter(spp == "UTST")
```

#### Step 1: Define the research questions
#### "Do snout-to-vent length, sex, and vegetation zone at time of capture significantly predict if a lizard tail is recorded as whole?"

#### Step 2: Examine data and possible correlations

```{r}
# Raw distributions
ggplot(sb_lizard, aes(x = SV_length))+
  geom_histogram()
ggplot(sb_lizard, aes(x = sex))+
  geom_bar()
ggplot(sb_lizard, aes(x = zone))+
  geom_bar()
ggplot(sb_lizard, aes(x = tail))+
  geom_bar()

# Relationships with predictor variables
ggplot(sb_lizard, aes(x = sex, y = tail))+
  geom_point()
ggplot(sb_lizard, aes(x = zone, y = tail))+
  geom_point()
ggplot(sb_lizard, aes(x = SV_length, y = tail))+
  geom_point()
```

#### Step 3 - Fit regression model

```{r}
# Set tail to be a factor with a reference level of "W"
sb_lizard$tail <- factor(sb_lizard$tail,
                         levels = c("W", "B"))

# Set sex to be a factor with a reference level of "F"
sb_lizard$sex <- factor(sb_lizard$sex,
                        levels = c("F", "J", "M"))

# Set zone to be a factor with a reference level of "C"
sb_lizard$zone <- factor(sb_lizard$zone,
                         levels = c("C", "G", "M", "T"))

# Logistic regression model
lizard_mod <- glm(tail ~ sex + zone + SV_length,
                  data = sb_lizard,
                  family = "binomial")
```

#### Step 4 - Evaluate model diagnostics

```{r}
summary(lizard_mod)
simulateResiduals(lizard_mod) %>% plot()
```

#### Step 5 - Interpret model results
#### The results of the logisitic regression indicate that sex did not have significantly different log-odds. The log-odds of tail wholeness increased significantly with grassland (B = 0.565, p = 0.001) and tarbush shrubland (B = 0.404, p = 0.024) in comparison with creosotebush shrubland. Snout-vent-length appeared to increase log-odds of tail wholeness as well (B = 0.045, p < 0.001). 

#### Predictive plots:
```{r}
# Simulate sex vector
sex_vector <- c(rep("F", 646),
                rep("J", 646),
                rep("M", 646))

# Simulate zone vector
zone_vector <- c(rep("C", 490),
                 rep("G", 490),
                 rep("M", 479),
                 rep("T", 479))

# Simulate SVL vector
SVL_vector <- rep(seq(from = 1, to = 102), 19)

# Join data
data_pred <- data.frame(sex_vector, zone_vector, SVL_vector)
colnames(data_pred) <- c("sex", "zone", "SV_length")

# Use original model to predict outcomes
prediction <- predict(lizard_mod,
                      newdata = data_pred,
                      type = "response",
                      se.fit = TRUE)

# Pull out all the predictions
data_fig <- data.frame(data_pred,
                       prediction$fit,
                       prediction$se.fit)

# Rename columns
colnames(data_fig) <- c("sex", "zone", "SV_length", "probability", "se")

# Graph the probabilities of tail wholeness
ggplot(data_fig, aes(x = SV_length,
                     y = probability))+
  geom_line(aes(color = sex))+
  geom_ribbon(aes(ymin = probability - se,
                  ymax = probability + se,
                  fill = sex), alpha = 0.3)+
  labs(x = "Snout-to-vent length (mm)", y = "Probability of a Whole Tail",
       color = "sex", fill = "sex")+
  theme_bw()

ggplot(data_fig, aes(x = SV_length,
                     y = probability))+
  geom_line(aes(color = zone))+
  geom_ribbon(aes(ymin = probability - se,
                  ymax = probability + se,
                  fill = zone), alpha = 0.3)+
  labs(x = "Snout-to-vent length (mm)", y = "Probability of a Whole Tail",
       color = "zone", fill = "zone")+
  theme_bw()
```

### (2) Poisson Regression

```{r}
# Load in data
npp_lizard <- read_csv("jrn_lizard_npp.csv")
```

#### Step 1: Define the research question
#### "Do season and plant species percent cover significantly predict lizard counts?

#### Step 2: Examine data and possible correlations

```{r}
# Raw counts
ggplot(npp_lizard, aes(x = lizard_count))+
  geom_histogram() # right skewed
ggplot(npp_lizard, aes(x = season))+
  geom_bar()
ggplot(npp_lizard, aes(x = BOER))+
  geom_histogram() # possible outlier
ggplot(npp_lizard, aes(x = LATR))+
  geom_histogram() # possible outlier
ggplot(npp_lizard, aes(x = PRGL))+
  geom_histogram() # possible outlier
ggplot(npp_lizard, aes(x = SCBR))+
  geom_histogram() # right skewed

# Relationships with predictor variables 
ggplot(npp_lizard, aes(x = season, y = lizard_count))+
  geom_point() # less in summer
ggplot(npp_lizard, aes(x = BOER, y = lizard_count))+
  geom_point() # positive linear
ggplot(npp_lizard, aes(x = LATR, y = lizard_count))+
  geom_point() # positive linear
ggplot(npp_lizard, aes(x = PRGL, y = lizard_count))+
  geom_point() # positive linear
ggplot(npp_lizard, aes(x = SCBR, y = lizard_count))+
  geom_point() # positive linear

# It looks like lizards are less common in the summer; include BOER and SCBR,
# may have stronger influence on counts
```

#### Step 3: Fit regression model

```{r}
# Set season as a factor with a reference level of "F"
npp_lizard$season <- factor(npp_lizard$season,
                            levels = c("F", "S", "W"))

# Poisson regression model
npp_lizard_mod <- glm(lizard_count ~ season + BOER + SCBR,
                      data = npp_lizard,
                      family = "poisson")
```

#### Step 4: Evaluate model diagnostics

```{r}
summary(npp_lizard_mod)
plot(npp_lizard_mod)

# Outliers include 50, 47, 4

# Remove outliers
npp_lizard2 <- npp_lizard[-c(4, 47, 50),]

# Refit the model
npp_lizard_mod2 <- glm(lizard_count ~ season + BOER + SCBR,
                      data = npp_lizard2,
                      family = "poisson")

# Re-examine model results
summary(npp_lizard_mod2)
plot(npp_lizard_mod2)

# Remove 1 more outlier
npp_lizard3 <- npp_lizard2[-1,]

# Refit the model
npp_lizard_mod3 <- glm(lizard_count ~ season + BOER + SCBR,
                      data = npp_lizard3,
                      family = "poisson")

# Re-examine model results
summary(npp_lizard_mod3)
plot(npp_lizard_mod3)
```

#### Step 5: Communicate Results

```{r}
# Formate model results as gt table
tbl_regression(npp_lizard_mod3)
```

#### Results: The results of the poisson regression suggest that the summer season has a strong negative effect on lizard counts (B = -0.81, p < 0.0001) when compared to fall while winter has no strong effect (B = 0.001, p = 0.99). Black grama grass had a strong negative influence on lizard counts (B = -0.087, p = 0.004) while burrograss had a strong positive influence (B = 0.50, p < 0.001). Note, the coefficents are for estimation of log-transformed counts and 4 outlier points we removed because they fell outside Cook's distance. 

### (3) See Github - [nholsclaw](https://github.com/nholsclaw/nholsclaw)
